Planning to Save Energy: How Information Format Affects Accuracy

### 2024-12-07 manuscript/test.qmd
update results section. And some formating improvements.

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
format:
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    html: 
      echo: true
      code-fold: show
      output-file: manuscript-code.html
    hugo-md:
        include: false
        echo: false
        html-math-method: mathjax
        output-file: manuscript_hugo.md
    gfm:
        echo: true
        output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
format-links:   
    - text: HTML
      href: manuscript.html
      icon: file-minus
    - text: HTML (with code)
      href: manuscript-code.html
      icon: file-plus
    - format: hikmah-manuscript-pdf
      text: PDF version
      icon: file-pdf
    - hikmah-manuscript-docx
    - gfm
date: 2024-12-07
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

For each family scenario, the particpipants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh). For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task). Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S. dollars. In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.

![Example trial in the energy planning task. Participants are shown the prior year electricity use of a household, and are tasked with creating a plan for the next year that will meet the energy reduction goal. Study 1 manipulates the format of the reduction goal to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}

@@

## Results

```{r}
pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

```

```{r}
#| label: tbl-s1-agg
#| tbl-cap: "Study 1: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# proportion matching vs close match

# proportion matching vs close match

# overall pct of subjects who matched their goal
s1_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Deviation From Goal' = median(mean_pct_change),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

 #pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal overall")

```

For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error. The current analysis employs a cumulative ordinal regression model, implemented via a Bayesian hierarchical framework [@burknerBrmsPackageBayesian2017] . This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category. All analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].

```{r}
#| label: tbl-s1-prop
#| tbl-cap: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

# compute percentage of subjects per accuracy level per group
observed_props_s1 <- s1_agg |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()

observed_props_s1 |> 
  mutate(n_total=sum(n)/4) |> 
  group_by(accuracy_level) |>
  mutate(ns=sum(n)/4) |> 
  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
  pivot_wider(
    names_from = 'Reference Class',
    values_from = c('% in Group')
  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

```

see @fig-s1-plot

```{r}
#| label: fig-s1-plot
#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

prop_acc_s1 <- s1_agg %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(accuracy_level) |>
    mutate(Probability = count / sum(count)) %>%
    ungroup()

ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "S1. % of Participants within each Accuracy Bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    theme_minimal() 
```

We analyzed planning accuracy using Bayesian ordinal regression. The dependent variable, plan error, was computed by binning the goal deviation into four ordered levels: exact match (0% error), minor deviations (0.01-2% error), moderate deviations (2.01-15% error), and major deviations (>15% error). For each comparison, we provide posterior odds ratios (OR) and their 95% CIs.

```{r}
#| label: tbl-s1-ord
##| tbl-cap: "Study 1: Ordinal Regression Model Results."

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 5000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)

#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

# as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, Category=.category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

#plot(conditional_effects(ordinal_model_s1),points=TRUE)

# or_table |>
#       rownames_to_column(var = "Term") |>
#       filter(!stringr::str_detect(Term, "Intercept")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))

# or_table <- data.frame(exp(fixef(ordinal_model_s1)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")
```

::: {#tbl-s1-or}
|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-1.94    |-3.19    |-0.74    |1.00 |
|Intercept[2]       |0.20     |-1.02    |1.42     |0.63 |
|Intercept[3]       |4.36     |3.11     |5.65     |1.00 |
|refClassPercentage |0.85     |-0.54    |2.23     |0.88 |
|refClassUSD        |2.72     |1.41     |4.06     |1.00 |

**Experiment 1**: Ordinal Regression results.

:::

::: {#tbl-s1-epred}
|Term               |Estimate |Est.Error |Q2.5 |Q97.5 |
|:------------------|:--------|:---------|:----|:-----|
|refClassPercentage |2.3      |2         |0.58 |9.3   |
|refClassUSD        |15.2     |2         |4.11 |58.1  |

**Experiment 1**: Odds ratios for group comparisons.

:::

The ordinal model is parameterized with thresholds (intercepts), and positive coefficients can indicate that it is more difficult to achieve higher accuracy categories in the USD condition. 
The model output suggests that, compared to the kWh condition, the USD condition shows a positive coefficient (Estimate = 2.72, 95% CI: 1.41 to 4.06) for the ordinal outcome. At least to me, this positive coefficient appears to indicate that, relative to the kWh reference class, participants in the USD condition are more likely to fall into higher numerical categories of the dependent variable coding. However, because the dependent variable is ordered from best (Exact match) to worst (Over 15% error), care is needed in interpretation. The Percentage condition coefficient (Estimate = 0.85, 95% CI: -0.54 to 2.23) is more uncertain, with its credible interval overlapping zero.  Posterior predictive checks (Figure 3) showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-plot2).

```{r}
#| label: fig-s1-plot2
#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
#| fig.width: 11
#| fig.height: 7

#bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])

# Plot predicted probabilities
# ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = Category)) +
#     geom_bar(stat = "identity", position = position_dodge()) +
#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
#                     position = position_dodge(width = 0.9), width = 0.2) +
#     labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
#     ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
#     theme_minimal()

#pp_check(ordinal_model_s1) + ggtitle("Study 1: Posterior Predictive Checks for Ordinal Regression Model") + theme_minimal()

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

# Experiment 2

The experimental procedures in study 2 are quite similar to those in study 1, but we also included a rounding manipulation (rounded vs. not rounded), and a manipulation of the goal (10% reduction vs. 15% rediction). We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, leaving a final sample of 196 participants.

Note that reference class remains a between-subjects variable, while percent goal, rounding, and state are within-subjects variables. In study 2, the new design is a 4 state temperature (2 warm vs. 2 cold states) X 2 task goal (10% vs. 15%) X 2 last year’s usage for the family and the state average (exact vs. rounded numbers) within X 3 task reference class (USD vs. Percentage vs. kWh) between.

## Results

```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.001)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# overall pct of subjects who matched their goal
s2_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Abs. Deviation From Goal' = median(mean_pct_change),
    'Log Deviation' = (median(mean_log_abs_error)),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

# s2_agg1 |> group_by(id,refClass,calc) |> 
#     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
#     summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
#     mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
#     n_accuracy=first(n_accuracy))

# s2_agg4 %>% ggplot(aes(x=refClass,y=mean_log_abs_error,fill=refClass)) + 
#   stat_summary(fun=mean, geom="bar") +
#   stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
#   geom_jitter(alpha = 0.2, width = 0.2, height = 0)

# s2_agg4 |> 
#   ggplot(aes(x = mean_log_abs_error,fill = refClass)) +
#   geom_density(alpha = 0.7)

# library(ggridges)
# s2_agg1 |>  ggplot(aes(y = refClass, x=log_abs_error,fill=refClass,col=refClass)) +
#     geom_density_ridges(alpha = 0.5) +
#     geom_vline(aes(xintercept=(log_abs_error)))

```

```{r}
#| label: tbl-s2-prop
#| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

# compute percentage of subjects per accuracy level per group
observed_props_s2 <- s2_agg1 |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()

observed_props_s2 |> 
  mutate(n_total=sum(n)/4) |> 
  group_by(accuracy_level) |>
  mutate(ns=sum(n)/4) |> 
  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
  pivot_wider(
    names_from = 'Reference Class',
    values_from = c('% in Group')
  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

```

```{r}
#| label: fig-s2-plot
#| tbl-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

# % of entire sample 
prop_combo_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "S1. % of Participants within each Accuracy Bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    theme_minimal() 
```

```{r}
#| label: fig-s2-plot2
#| fig-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
#| fig.width: 11
#| fig.height: 9

# % of entire sample 
prop_combo_s2_full <- s2_agg1 %>%
    group_by(refClass, accuracy_level, pct_goal,rounded) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

ggplot(prop_combo_s2_full, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "S1. % of Participants within each Accuracy Bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    facet_wrap(~pct_goal*rounded) +
    theme_minimal() 
```

```{r}
#| label: tbl-s2-ord
##| tbl-cap: "Study 2: Ordinal Regression Model Results."

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
#summary(ordinal_model_s2_logit)

# as.data.frame(describe_posterior(ordinal_model_s2_logit, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
#   kable(escape = FALSE, booktabs = TRUE, align = c("l"), row.names = FALSE)

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
 # group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
  group_by("Reference Class"=refClass, Category=.category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

#pred_summary_s2 |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)

# Create a new data frame with the predictor values
# new_data <- expand.grid(
#   refClass = c("kWh", "Percentage", "USD"),
#   id = unique(s2_agg1$id)[1],      # Take first id
#   state = unique(s2_agg1$state)[1], # Take first state,
#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
# )

# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")

```

::: {#tbl-s2-or}
|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-2.13    |-3.39    |-0.86    |1.00 |
|Intercept[2]       |-0.62    |-1.89    |0.63     |0.84 |
|Intercept[3]       |3.15     |1.88     |4.42     |1.00 |
|refClassPercentage |0.83     |-0.64    |2.33     |0.87 |
|refClassUSD        |1.87     |0.31     |3.35     |0.99 |
|roundedRounded     |-0.66    |-1.01    |-0.31    |1.00 |
|pct_goal15%        |-0.44    |-0.79    |-0.10    |0.99 |

**Experiment 2.** Ordinal Regression Model Results.

:::

::: {#tbl-s2-epred}
|comparison           | odds_ratio| ci_lower| ci_upper|
|:--------------------|----------:|--------:|--------:|
|Percentage vs kWh    |       3.02|     0.53|    10.31|
|USD vs kWh           |       8.80|     1.37|    28.42|
|Rounded vs Not       |       0.53|     0.36|     0.73|
|15% Goal vs 10% Goal |       0.66|     0.45|     0.91|

**Experiment 2.** Odds ratios for group comparisons.

:::

```{r}
#| label: fig-s2-plot3

# ggplot(pred_summary_s2, 
#         aes(x = .category, y = mean_prob, fill = refClass )) +
#     geom_col(position = position_dodge(width = 0.8)) +
#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
#                     position = position_dodge(width = 0.8)) +
#     labs(x = "Reference Class", 
#         y = "Predicted Probability",
#         fill = "Accuracy Level",
#         title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
#     scale_fill_brewer(palette = "RdYlBu") +
#     theme_minimal()

pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

```

### Individual Differences

see @fig-s2-indv

```{r}
#| label: fig-s2-indv
#| fig-cap: "Study 2: Individual performance in the energy planning task, colored by reference class. The dashed line represents the target reduction goal. Participants are shown along the y axis, those who fall above or below the dashed line have not met the target goal.  The x-axis represents the percent change in energy usage from the prior year."
#| fig.width: 11
#| fig.height: 9

s2_agg1 |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
    filter(plan=="plan1",rounded=="Rounded") |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  filter(mean_abs_error <= 0.50) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y=element_text(face = "plain", size = rel(0.7))) + 
    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
                    labels = scales::percent_format(accuracy = 1))

```

```{r}
#| label: fig-s2-indv2
#| fig-cap: "Study 2: Respones patterns for a subset of individiual participants. Black points are participant responses, red points are the state average, and blue points are the family average. The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
#| fig.width: 11
#| fig.height: 10

s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discusion

Karjalainen 2011 - people prefer information about price [@karjalainenConsumerPreferencesFeedback2011]

# References
::: {#refs}
:::

>>>>

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
author:
  - name: Thomas E. Gorman
    url: https://tegorman13.github.io/
    orcid: "0000-0001-5366-5442"
    affiliations: 
      - ref: cclab
      - ref: clara
    attributes: 
      corresponding: true
  - name: Torsten Reimer
    url: https://web.ics.purdue.edu/~treimer/
    orcid: 0000-0002-7419-0076
    affiliations:
      - ref: cclab
      - ref: clara
  - name: Juan Pablo Loaiza Ramirez
    url: https://www.cla.purdue.edu/directory/profiles/loaiza-ramirez,juan-pablo.html
    orcid: 0000-0001-9663-0522
    affiliations:
      - ref: cclab
      - ref: clara
  - name: Hayden Barber
    url: https://www.sdstate.edu/directory/hayden-barber
    orcid: 0000-0002-3465-8615
    affiliations:
      - ref: sdsu
affiliations:
  - id: cclab
    name: Communication and Cognition Lab, Purdue University
    city: West Lafayette
    region: IN
    country: USA
    postal-code: 47907
    url: https://web.ics.purdue.edu/~treimer/
  - id: clara
    name: College of Liberal Arts Research Academy
    department: Purdue University
    city: West Lafayette
    region: IN
    country: USA
    postal-code: 47907
    url: https://cla.purdue.edu/about/college-initiatives/research-academy/
  - id: sdsu
    name: School of Communication & Journalism, South Dakota State University
    city: Brookings
    region: SD
    country: USA
    url: https://www.sdstate.edu/directory/hayden-barber
format:
    # html: 
    #     grid:
    #         sidebar-width: 220px
    #         body-width: 1200px
    #         margin-width: 170px
    #         gutter-width: 1.0rem
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        code-fold: true
        code-summary: "Show the code"
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    hugo-md:
        include: true
        echo: false
        html-math-method: mathjax
        output-file: test_hugo.md
    gfm:
        echo: true
        output-file: test_gfm.md
    hikmah-manuscript-pdf:
        output-file: test.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
# format-links:   
#     - html
#     - gfm
#     - format: hikmah-manuscript-pdf
#       text: PDF version
#       icon: file-pdf
date: 2024-12-03
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

@@
